Lawvere-Tierney Sheafification\\ in Homotopy Type Theory

### 2015-12-17 paper/paper_acm.tex
reduce again

\category{CR-number}{subcategory}{third-level}

@@

In this paper, we use the double negation modality at level $-1$ only,
and then extend it to higher homotopy types using sheafification. This
way, we can add classical reasoning in the type theory without
collapsing the theory to mere propositions.

@@

As $\modal_{n+1} T$ is contractible, any path space of $\modal_{n+1} T$ is
contractible, in particular $(\nu_T a=\nu_T b)$, which proves left
exactness.

@@

One issue with universe polymorphism lies in the management of
recursive definitions. Indeed, the following recursive definition of
sheafification
%
\[ \begin{array}{l}
   \modal : \forall \ (n : nat), \ \Type_n \to \Type_n 
   \\
    \modal_{-1\phantom{n}}(T) \defeq\lnot\lnot T \\
      \displaystyle{\modal_{n+1}(T)} \defeq  
      \displaystyle{\sum_{u:T \to \Type_n^\modal} \!\!\!\!\modal_{-1} 
      \left\|
      \sum_{a:T} u= (\lambda t,~\modal_n (a=t))
      \right\|}
    \end{array}
\]
%
is not allowed. 
%
This is because Coq forces the universe of the first $\Type_n$
occurring in the definition to be the same for every $n$, whereas the
universe of the first $\Type_{n+1}$ occurring in $\modal_{n+1}$ should be at
least one level higher as the one of $\Type_n$ occurring in
$\modal_{n}$ because of the use of $\Sigma$-type over
$T \to \Type_n^\modal$ and equality on the return type of $\modal_n$.
%
% \nt{explain why it is an issue.}
%
Thus, the induction step presented in this paper has been formalized,
but the complete recursive sheafification can not be defined for the
moment.
%
Note that the same increasing in the universe levels occurs in the
Rezk completion for categories~\cite{rezk}. In the definition of the
completion, they use the Yoneda embedding and representable functors,
which is similar to our use of characteristic functions.

@@

\section*{Acknowledgment}

We would like to thank all maintainers of the Coq/HoTT library, who
have provided a very solid base for our formalization. 
We also would like
to thank Bas Spitters for helpful discussions on the subject.

>>>>

\category{F}{8}{3}

@@

In this paper, we use the double negation modality at level $-1$ only,
and then extend it to higher homotopy types using sheafification. This
way, we can add classical reasoning in the type theory without
collapsing it to mere propositions.

@@

As $\modal_{n+1} T$ is contractible, path spaces of $\modal_{n+1} T$ are
contractible, in particular $(\nu_T a=\nu_T b)$, which proves left
exactness.

@@

One issue with universe polymorphism lies in the management of
recursive definitions. Indeed, the following recursive definition of
sheafification
%
\vspace{-0.05em}
\[ \begin{array}{l}
   \modal : \forall \ (n : nat), \ \Type_n \to \Type_n 
   \\
    \modal_{-1\phantom{n}}(T) \defeq\lnot\lnot T \\

      \displaystyle{\modal_{n+1}(T)} \defeq  
      \displaystyle{\sum_{u:T \to \Type_n^\modal} \!\!\!\!\modal_{-1} 
      \left\|
      \sum_{a:T} u= (\lambda t,~\modal_n (a=t))
      \right\|}
    \end{array}
\vspace{-0.3em}
\]
%
is not allowed. 
%
This is because Coq forces the universe of the first $\Type_n$
occurring in the definition to be the same for every $n$, whereas the
universe of the first $\Type_{n+1}$ occurring in $\modal_{n+1}$ should be at
least one level higher as the one of $\Type_n$ occurring in
$\modal_{n}$ because of the use of $\Sigma$-type over
$T \to \Type_n^\modal$ and equality on the return type of $\modal_n$.
%
% \nt{explain why it is an issue.}
%
Thus, the induction step presented in this paper has been formalized,
but the complete recursive sheafification can not be defined for the
moment.
%
Note that the same increasing in the universe levels occurs in the
Rezk completion for categories~\cite{rezk}. In the definition of the
completion, they use the Yoneda embedding and representable functors,
which is similar to our use of characteristic functions.

@@

% \section*{Acknowledgment}

% We would like to thank all maintainers of the Coq/HoTT library, who
% have provided a very solid base for our formalization. 
% We also would like
% to thank Bas Spitters for helpful discussions on the subject.
